---
title: "Untitled"
output: html_document
date: "2025-11-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)
library(here)
library(srvyr)
library(purrr)
library(forcats)


options(scipen = 999)

## harmonize poll data with the cnesus variables

data <- read_csv(here("Data and script", "data", "data", "resegmentation.csv"))




data <- data %>%
  mutate(
    ethnicity = case_when(
      ethnicity == "White (British/Irish/Other)" ~ "White",
      ethnicity == "Black/Black British" ~ "Black",
      ethnicity == "Asian/Asian British" ~ "Asian",
      ethnicity == "Mixed descent (e.g. White & Asian, White & Black)" ~ "Mixed",
      ethnicity == "Other" ~ "Other",
      TRUE ~ NA_character_
    ), 
    education = case_when(
     education %in% c(
    "University Education Completed (First Degree e.g. BA, BSc)",
    "Postgraduate Education Completed (e.g. Masters)",
    "Doctorate, Post-doctorate or equivalent (Higher Degree)"
  ) ~ "University",
    education %in% c(
    "Incomplete Secondary Education (Below GCSE/O Level)",
    "Secondary Education Completed (GCSE/O Level/CSE or equivalent)",  
    "Secondary Education Completed (A Level or equivalent)"
  ) ~ "Secondary",
     education %in% c(
    "Vocational or Technical Qualifications Completed (e.g. HND, NVQ)",
    "Some Vocational or Technical Qualifications"
  ) ~ "Vocational",
  education == "Prefer not to answer" ~ NA_character_,  # Added this
  TRUE ~ NA_character_
),
degree = case_when(education == "University" ~ "Degree", 
                   education %in% c("Secondary", "Vocational") ~ "No Degree", 
                   TRUE ~ NA_character_),
    seg = case_when(
      seg %in% c(
        "DE Semi or unskilled manual work",
        "C2 - Skilled manual worker (e.g. skilled bricklayer, carpenter, electrician, plumber, painter, bus/ambulance driver, HGV/train driver, AA patrolman, mechanic)"
      ) ~ "Semi or unskilled manual work",
      seg %in% c(
        "C1 - Supervisory or clerical/junior managerial/professional/administrative (e.g. office worker, student doctor, foreman with 25+ employees, salesperson, policeman, nurse, secretary, self-employed)",
        "B - Intermediate managerial/professional/administrative (e.g. newly qualified (under 3 years) doctor, solicitor, board director in a small organisation, middle manager in a large organisation, principal officer in the civil service/local government, teacher)",
        "Retired"
      ) ~ "Supervisory or clerical/junior managerial/professional/administrative",
      seg %in% c(
        "Intermediate managerial/professional/administrative",
        "A - Higher managerial/professional/administrative (e.g. established doctor, solicitor, board director in a large organisation (200+ employees), top level civil servant/public service employee)"
      ) ~ "Intermediate managerial/professional/administrative",
      seg == "Higher managerial/professional/administrative" ~ "Higher managerial/professional/administrative",
      seg == "Full time education/student" ~ "Full time education/Student",
      seg %in% c("Unemployed", "Looking after home or family", "Prefer not to say") ~ NA_character_,
      TRUE ~ NA_character_
    ), 
    religion = case_when(
      grepl("^Christian", religion) ~ "Christian", 
      grepl("Any other", religion) ~ "Other religion", 
      religion == "Prefer not to say" ~ "Not answered", 
      TRUE ~ religion
    ),  
    welshlang = case_when(
      welshlang == "I can speak Welsh" ~ "Can speak Welsh", 
      welshlang %in% c("I cannot speak Welsh but I can understand, read or write Welsh", 
                       "I can't speak, understand, read or write Welsh") ~ "Cannot speak Welsh", 
      TRUE ~ NA_character_
    )
  )

##data_small <- data %>%
##  select(borninuk, cid, seg, region, ethnicity, age_grouped, religion, gender, degree, education, welshlang, UKNatRepWeight) %>%
##  filter(if_all(everything(), ~ !is.na(.))) %>% 
##  rename(sex = gender, 
  ##       age_group = age_grouped,
    ##     socio_econ = seg)



data_small <- data %>%
  select(borninuk, cid, seg, region, ethnicity, age_grouped, religion, gender, degree, education, degree, welshlang, UKNatRepWeight) %>%
  rename(sex = gender, 
         age_group = age_grouped,
         socio_econ = seg)


## wighting the data 

poll_svy <- data_small %>%
  as_survey_design(ids = 1, weights = UKNatRepWeight)

```



```{r}

## recode cnesus values to haromize with  poll data 

recode_census <- function(x, ...){
  #   if("degree" %in% names(x)){
  #   x <- x %>% 
  #     mutate(degree = case_when(
  #       degree == "No Degree" ~ "No degree", 
  #       TRUE ~ degree
  #     ))
  # }
  
    if("education" %in% names(x)){
      x <- x %>% 
      mutate(education = case_when(
        education %in% c(
          "Incomplete Secondary Education (Below GCSE /O Level)",
          "Secondary Education Completed (GCSE /O Level /CSE or equivalent)",
          "Secondary Education Completed (A Level or equivalent)"
        ) ~ "Secondary",
        education == "Vocational or Technical Qualifications Completed (e.g. HND, NVQ)" ~ "Vocational",
        education == "University Education Completed (First Degree e.g. BA, BSc)" ~ "University",
        TRUE ~ education  # Keep original value instead of NA
    ), 
    degree = case_when(
      education == "University" ~ "Degree", 
      education %in% c("Secondary", "Vocational") ~ "No Degree", 
      TRUE ~ NA_character_
    ))
    }
  if("socio_econ" %in% names(x)){
    x <- x %>% 
      mutate(socio_econ = case_when(
        socio_econ == "Intermediate managerial/ professional/ administrative"  ~ "Intermediate managerial/professional/administrative", 
        socio_econ == "Supervisory or clerical/ junior managerial/ professional/ administrative" ~ "Supervisory or clerical/junior managerial/professional/administrative",
        socio_econ == "Higher managerial/ professional/ administrative" ~ "Supervisory or clerical/junior managerial/professional/administrative", 
        socio_econ == "Skilled manual work" ~ "Semi or unskilled manual work", 
        TRUE ~ socio_econ)
      )
  }
  if("religion" %in% names(x)){
    x <- x %>% 
      mutate(religion = ifelse(religion == "Does not apply", NA, religion))
  }
  if("ctry_birth" %in% names(x)){
    x <- x %>% rename(borninuk = ctry_birth) %>% 
      mutate(borninuk = ifelse(borninuk == "Born in UK", "Yes", "No"))
  }
  if("sex" %in% names(x)){
    x <- x %>% 
      mutate(sex = ifelse(sex == "Total", NA, sex) )  
  }
  if("region" %in% names(x)){
    x <- x %>% 
      mutate(region = case_when(
        region == "London" ~ "Greater London", 
        region == "North East" ~ "North East England", 
        region == "North West" ~ "North West England",
        region == "South East" ~ "South East England",
        region == "South West" ~ "South West England",
        region == "Yorkshire and The Humber" ~ "Yorkshire and the Humber", 
        TRUE ~ region
      ))
  }
na.omit(x)
}


```


Pre-process step for the census data that will have degree as well as education. 

```{r}
cens_files <- list.files("Data and script/all_census_data",
                         full.names = TRUE, recursive = TRUE)
for(f in cens_files){
  cens <- read_csv(f)
  cens <- recode_census(cens)
  if(all(c("degree", "education") %in% names(cens))){
    incl_vars <- intersect(names(cens), names(data_small))
    incl_vars <- setdiff(incl_vars,"education")
    
    # Add this: rename percent to proportion if it exists
    if("percent" %in% names(cens) && !"proportion" %in% names(cens)){
      cens <- cens %>% rename(proportion = percent)
    }
    
    cens_degree <- cens %>% 
      group_by(across(all_of(incl_vars))) %>% 
      summarise(across(c(n, proportion), sum))  # Now just sum n and proportion
    newf <- gsub("edu[cation]*", "degree", f)
    write_csv(cens_degree, newf, append = FALSE)
  }
}

```




```{r}
## takes poll data and census file names and computes comparison of disitburiotns - for census and poll data

compare_distributions <- function(sample_data, census_file, write=TRUE, output_file = NULL, output_dir = "output", ...){
  
  cens <- read_csv(census_file)
  cens <- recode_census(cens)
  
  # Rename 'percent' column to 'proportion' for consistency if it exists
  if("percent" %in% names(cens) && !"proportion" %in% names(cens)){
    cens <- cens %>% rename(proportion = percent)
  } else if("percent" %in% names(cens) && "proportion" %in% names(cens)){
    cens <- cens %>% select(-proportion) %>% rename(proportion = percent)
  }
  
  if(all(c("education", "degree") %in% names(cens))){
    cens <- cens %>% select(-degree)
  }
  if(inherits(sample_data, "data.frame")){
    incl_vars <- intersect(names(cens), names(sample_data))
  }else{
    incl_vars <- intersect(names(cens), names(sample_data$variables))
  }
  cens <- cens %>% 
  group_by(across(all_of(incl_vars))) %>%
  summarise(proportion = sum(proportion))
  
  
  ## if data frame it is unweighted 
  if(inherits(sample_data, "data.frame")){
    samp <- sample_data %>% 
      group_by(across(all_of(incl_vars))) %>% 
      tally() %>% 
     #  filter(!if_any(all_of(incl_vars), is.na)) %>%  ## check if this is right way to deal with NA 
      ungroup() %>% 
      mutate(sample_proportion = n/sum(n)) %>% 
      select(-n)
    ## otherwise get the weighted size 
  }else{
    samp <- sample_data %>% 
      group_by(across(all_of(incl_vars))) %>% 
      survey_tally() %>% 
      select(-n_se) %>% 
    # filter(!if_any(all_of(incl_vars), is.na)) %>%  ## this as well 
      ungroup() %>% 
      mutate(sample_proportion = n/sum(n)) %>% 
      select(-n)
  }
  
  out <- left_join(cens, samp) %>% 
    mutate(sample_proportion = ifelse(is.na(sample_proportion), 0, sample_proportion),
           # proportion is already in percent, just round it
           proportion = round(proportion, 2),
           # sample_proportion needs to be converted to percent
           sample_proportion = round(sample_proportion * 100, 2),
           difference = proportion - sample_proportion) %>% 
    arrange(desc(abs(difference))) %>% 
    rename(census = proportion, 
           poll = sample_proportion)
  
  if(write){
    if(is.null(output_file)){
      fname <- if(grepl("/", census_file, fixed=TRUE)){
        tail(strsplit(census_file, "/", fixed=TRUE)[[1]], n=1)
      } else {
        census_file
      }
      fname <- paste0(output_dir, "/compare_", fname)
    }else{
      fname <- paste0(output_dir, "/", output_file)
    }
    if(!dir.exists(output_dir)){
      message("Creating output directory.\n")
      dir.create(output_dir)
    }
    write_csv(out, fname, append = FALSE)
  }
  return(out)
}


```


## Considering Welsh Language

```{r}
tab <- table(data_small$age_group, data_small$welshlang)
# number of observations by age group
n_age <- rowSums(tab)
# simulated number of welsh speakers based on sample proportion by age group
sim_n <- lapply(1:10000, \(i)rbinom(length(n_age), n_age, prob = tab[,1]/n_age))
## make list into matrix
sim_n <- do.call(rbind, sim_n)

## simulated proportion of welsh speakers by age group
sim_p <- apply(sim_n, 1, \(x)x/n_age)
## 95% lower and upper bounds of proportion of welsh speakers by age group
cbind(apply(sim_p, 1, \(x)quantile(x, c(.025))), 
apply(sim_p, 1, \(x)quantile(x, c(.975))))

```

The range on these values is around 15-20 percentage points indicating that the estimate of the proportion of Welsh language speakers from the polling data is really imprecise.  This suggests that there could be meaningful departures of the sample values from the census values simply due to sampling variability alone. 


```{r}


cens_files <- list.files(here("Data and script", "all_census_data"), 
                         full.names = TRUE, recursive = TRUE)

for(f in cens_files){
  cens <- read_csv(f)
  cens <- recode_census(cens)
  if(all(c("degree", "education") %in% names(cens))){
    incl_vars <- intersect(names(cens), names(data_small))
    incl_vars <- setdiff(incl_vars,"education")
    
    if("percent" %in% names(cens) && !"proportion" %in% names(cens)){
      cens <- cens %>% rename(proportion = percent)
    }
    
    cens_degree <- cens %>% 
      group_by(across(all_of(incl_vars))) %>% 
      summarise(across(c("n", "proportion"), sum))
    
    newf <- gsub("edu[cation]*", "degree", f)
    write_csv(cens_degree, newf, append = FALSE)
  }
}

# Load census files
cens_files <- list.files(here("Data and script", "all_census_data"), 
                         full.names = TRUE, recursive = TRUE)

## Compare each file to data_small - the unweighted data 
raw_comparisons <- lapply(cens_files, \(x)compare_distributions(data_small, x, write=TRUE))

## Compare census data to poll_svy - the weighted data
wt_comparisons <- lapply(cens_files, \(x)compare_distributions(poll_svy, x, write=TRUE, output_dir = "wt_output"))
wt_comparisons <- lapply(wt_comparisons, \(x)x %>% rename(wt_poll = poll, wt_difference = difference))

combined_comparisons <- lapply(seq_along(raw_comparisons), \(i){
  left_join(raw_comparisons[[i]], wt_comparisons[[i]] %>% select(-census))
})

## How often does the weighting decrease the absolute difference i.e. improve census match 
out_comp <- lapply(combined_comparisons, \(x)x %>% ungroup %>% 
         summarise(wt_better_sum = sum(abs(wt_difference) < abs(difference)), 
                   n_cells = n(), 
                   wt_better_prop = mean(abs(wt_difference) < abs(difference))))
names(out_comp) <- gsub(".*combined_(.*)\\.csv", "\\1", cens_files)
out_comp

# Create output directories if they don't exist
if(!dir.exists("output/wt_comparisons")){
  dir.create("output/wt_comparisons", recursive = TRUE)
}
if(!dir.exists("output/raw_comparisons")){
  dir.create("output/raw_comparisons", recursive = TRUE)
}

# MODIFIED: Extract demographic names from census filenames
get_demo_name <- function(filepath) {
  fname <- basename(filepath)
  # Remove "combined_" prefix and ".csv" suffix
  demo_name <- str_remove(fname, "^combined_")
  demo_name <- str_remove(demo_name, "\\.csv$")
  return(demo_name)
}

# Get demographic names for each file
demo_names <- sapply(cens_files, get_demo_name)

# Write weighted comparisons with demographic names
for (i in seq_along(wt_comparisons)) {
  write_csv(wt_comparisons[[i]], 
            file = paste0("output/wt_comparisons/", demo_names[i], ".csv"))
}

# Write raw comparisons with demographic names
for (i in seq_along(raw_comparisons)) {
  write_csv(raw_comparisons[[i]], 
            file = paste0("output/raw_comparisons/", demo_names[i], ".csv"))
}

out_comp_df <- bind_rows(out_comp, .id = "breakdown")
write_csv(out_comp_df, here("Data and Script", "output", "weighting_performance_summary.csv"))

out_comp_df





```




```{r}


library(tidyverse)
library(here)
library(purrr)
library(fs)

unw_dir <- here("output", "raw_comparisons")  
wt_dir  <- here("output", "wt_comparisons")   

read_and_join_pair <- function(fname) {
  unw <- read_csv(file.path(unw_dir, fname), show_col_types = FALSE) |>
    rename(diff_unw = difference, poll_unw = poll)

  wt  <- read_csv(file.path(wt_dir, fname), show_col_types = FALSE) |>
    rename(diff_wt = wt_difference, poll_wt = wt_poll)

  key_cols <- setdiff(names(unw), c("census", "poll_unw", "diff_unw"))

  joined <- unw |>
    select(all_of(key_cols), census, poll_unw, diff_unw) |>
    inner_join(
      wt |> select(all_of(key_cols), poll_wt, diff_wt, census),
      by = key_cols
    ) |>
    select(-census.y) |>
    rename(census = census.x)

  joined
}

file_list  <- list.files(unw_dir, pattern = "\\.csv$", full.names = FALSE)
joint_list <- map(file_list, read_and_join_pair)
names(joint_list) <- tools::file_path_sans_ext(file_list)

joined_dir <- here("output", "joined_comparisons")
if (!dir_exists(joined_dir)) dir_create(joined_dir)

walk2(
  joint_list,
  names(joint_list),
  ~ write_csv(.x, file = file.path(joined_dir, paste0(.y, ".csv")))
)

# Function to add cell-level bias reduction stats
add_cell_bias_stats <- function(fname) {
  filepath <- file.path(joined_dir, fname)
  df <- read_csv(filepath, show_col_types = FALSE)
  
  df_with_stats <- df %>%
    mutate(
      abs_diff_unw = round(abs(diff_unw), 2),
      abs_diff_wt = round(abs(diff_wt), 2),
      bias_reduction = round(abs_diff_unw - abs_diff_wt, 2),
      bias_red_pct = round(100 * (abs_diff_unw - abs_diff_wt) / abs_diff_unw, 2)
    )
  
  write_csv(df_with_stats, filepath)
  cat("Updated:", fname, "\n")
  
  return(df_with_stats)
}

joined_files <- list.files(joined_dir, pattern = "\\.csv$", full.names = FALSE)
updated_files <- map(joined_files, add_cell_bias_stats)

cat("\nAll files updated with cell-level bias statistics!\n")

# Summarise bias with clean demographic names
summarise_bias <- function(fname) {
  df <- read_csv(file.path(joined_dir, fname), show_col_types = FALSE)

  MAB_unw <- mean(abs(df$diff_unw), na.rm = TRUE)
  MAB_wt  <- mean(abs(df$diff_wt),  na.rm = TRUE)
  MAB_unw_w <- sum(abs(df$diff_unw)*(df$census/100), na.rm = TRUE)
  MAB_wt_w  <- sum(abs(df$diff_wt)*(df$census/100),  na.rm = TRUE)

  # Just keep the filename as-is (with underscores)
  joint_distribution <- tools::file_path_sans_ext(fname)

  tibble(
    joint_distribution = joint_distribution,
    MAB_unw        = round(MAB_unw, 3),
    MAB_wt         = round(MAB_wt, 3),
    MAB_reduction  = round(MAB_unw - MAB_wt, 3),
    MAB_red_pct    = round(100 * (1 - MAB_wt / MAB_unw), 2),
    MAB_unw_w        = round(MAB_unw_w, 3),
    MAB_wt_w         = round(MAB_wt_w, 3),
    MAB_reduction_w  = round(MAB_unw_w - MAB_wt_w, 3),
    MAB_red_pct_w    = round(100 * (1 - MAB_wt_w / MAB_unw_w), 2)
  )
}

summarise_bias <- function(fname) {
  df <- read_csv(file.path(joined_dir, fname), show_col_types = FALSE)

  MAB_unw <- mean(abs(df$diff_unw), na.rm = TRUE)
  MAB_wt  <- mean(abs(df$diff_wt),  na.rm = TRUE)
  MAB_unw_w <- sum(abs(df$diff_unw)*(df$census/100), na.rm = TRUE)
  MAB_wt_w  <- sum(abs(df$diff_wt)*(df$census/100),  na.rm = TRUE)

  # Extract demographic name
  joint_name <- tools::file_path_sans_ext(fname)
  
  # Option 1: Use lowercase 'x' instead of special character
  joint_distribution <- str_replace_all(joint_name, "_", " x ")
  
  # Option 2: Keep underscores
  # joint_distribution <- joint_name

  tibble(
    joint_distribution = joint_distribution,
    MAB_unw        = round(MAB_unw, 3),
    MAB_wt         = round(MAB_wt, 3),
    MAB_reduction  = round(MAB_unw - MAB_wt, 3),
    MAB_red_pct    = round(100 * (1 - MAB_wt / MAB_unw), 2),
    MAB_unw_w        = round(MAB_unw_w, 3),
    MAB_wt_w         = round(MAB_wt_w, 3),
    MAB_reduction_w  = round(MAB_unw_w - MAB_wt_w, 3),
    MAB_red_pct_w    = round(100 * (1 - MAB_wt_w / MAB_unw_w), 2)
  )
}

bias_summary <- map_dfr(joined_files, summarise_bias)
bias_summary <- bias_summary %>% arrange(desc(MAB_reduction_w))

print(bias_summary)


write_csv(bias_summary, here("output", "bias_summary_with_demographics.csv"))

write_csv(out_comp_df, here("output", "weighting_performance_summary.csv"))

```



## First four columns are straight averages that do not weight by cell size
## last four columns are the same quantities, but weighted by cell size
## if MAB_unw >> MAB_unw_w, then the errors are largely coming from smaller cells
## if MAB_unw << MAB_unw_w, then the errors are largely coming from larger cells. 
## same logic holds for MAB_wt and MAB_wt_w. 


```




```


```{r}
##The name of the joint distribution the row refers to (e.g. age_ethnicity, religion_education).

##MAB_unw

#Mean Absolute Bias for the UNweighted data.

#It is the average of |diff_unw| across all cells in that joint (e.g. all age × ethnicity combinations).

#Interpreted as: “On average, unweighted cell proportions are this many percentage points away from the census.”​

#MAB_wt

#Mean Absolute Bias for the weighted data.

#It is the average of |diff_wt| across all cells in that joint.

#Interpreted as: “On average, weighted cell proportions are this many percentage points away from the census.”​

#MAB_reduction

#Absolute reduction in mean absolute bias from weighting.

#Computed as MAB_unw − MAB_wt.
#Interpreted as: “Weighting reduced average bias by this many percentage points.”​
#Positive = weighting reduced average bias; negative = weighting increased average bias.​

#MAB_red_pct

write.csv(bias_summary, here("output", "bias_summary.csv"), row.names = FALSE)

#Percentage reduction in mean absolute bias.

#Computed as 100 × (1 − MAB_wt / MAB_unw).

#Positive values: percentage improvement; negative values: percentage worsening of bias after weighting.​



## test 

data_small %>%
  summarise(
    percentage = mean(education == "Secondary" & sex == "Female" & age_group == "75+", na.rm = TRUE) * 100,
    count = sum(education == "Secondary" & sex == "Female" & age_group == "75+", na.rm = TRUE),
    total = n()
  )



data_small %>%
  summarise(
    percentage = mean(education == "Secondary" & sex == "Male" & age_group == "25-34", na.rm = TRUE) * 100,
    count = sum(education == "Secondary" & sex == "Male" & age_group == "25-34", na.rm = TRUE),
    total = n()
  )



data_small %>%
  summarise(
    percentage = mean(education == "Vocational" & sex == "Female" & age_group == "55-64", na.rm = TRUE) * 100,
    count = sum(education == "Vocational" & sex == "Female" & age_group == "55-64", na.rm = TRUE),
    total = n()
  )

```